\chapter{Background}

The introductory material we provide in this section sets up the scene for the work that will be later explained. It follows a path that goes from the early days of program reasoning all the way to modern techniques that allow to deal with concurrency. On top of this, we look at database theory from a transactional perspective, analyzing concurrency control paradigms and consistency models.

\section{History of Concurrent Reasoning}

Before entering the world of modern program logics for concurrency, we study the building blocks that serve as its foundation. First and foremost, the work done by Tony Hoare provides us with the starting point for reasoning about sequential programs using an axiomatic approach. The latter is expanded in order to cope with concurrency, and later enhanced through a methodology allowing the abstraction of effects that concurrent threads may have on the program state.

\input{./david/background/history.tex}

\section{Modern Concurrent Reasoning}

The explanation of logic reasoning techniques continues by considering approaches that are currently used in concurrent systems. These frameworks allow to tackle the verification of heap-manipulating programs thanks to the advances brought by separation logic. Key results in this area have been achieved through the use of abstraction in specifications, which allows to use program logics in a client-module setting, and the formalisation of atomicity, a fundamental concept in concurrency where an operation is seen as happening in one single step.

\input{./david/background/logic.tex}

\section{Database Theory}

This section shifts the focus from program logics to the main theory behind a database. This is generically abstracted from a particular instantiation, which enables us to describe its properties from a high-level point of view, without considering its implementation. We are mainly interested in the concurrent aspects of modern databases, therefore we concentrate on transactions and the way they are managed.

\input{./david/background/transactions.tex}