\subsubsection{Verification}
\paragraph{\bf Semantics the code}
Let \( \repl \in \Repls \) denotes the set of totally ordered replicates.
Each replicate can have multiple clients, and 
each clients can commit a sequence of either read-only transitions or single-write transactions.
To model these, we annotate the transaction identifier with replicate \( \repl \), client \( \cl \), 
local time of the replicate \( n \) and read-only transactions counter \( n' \), \ie \( \txidCOPS{\repl}{\cl}{n}{n'} \).
Note that the \( (n, \repl, n') \) can be treated as a single number that \( n \) are the higher bits, 
\( \repl \) the middle bits and \( n' \) the lower bits.
For a new single-write transaction, it is allocated with a transaction identifier with larger local time,
and for a read-only transactions, it is allocated with a transaction identifier with larger read-only counter.
There is a total order among transitions from the same replica and from the same client.

To model the dependencies of each version,
We extend version from \cref{def:his_heap} 
with the set of all versions it dependencies on, \( \dep \in \pset{\Keys \times \TxID} \).
The function \( \depOf{\ver} \) denotes the dependencies set of the version.
We use \( \mkvsCOPS \) for key-value store whose versions contain the dependencies.

We use view to model the client context,
that is, a version is included in a context if and only if such version is in the view of the client.
We also use view to model a replica state,
that is, if a replica contains a version if and only if such version in the view of the replica.
For readability, we annotate view with either a replica, \( \viREPL \), or a client, \( \viCL \).
The view environment is extended with replicas and their views, \( \vienvCOPS : (\Repls \times \Clients ) \parfinfun \Views \).
We give the following semantics to capture the behaviours of the code.

\paragraph{\bf Write}
For purpose of verification, we eliminate code for performance, and put the client API and replica API in the same function (\cref{lst:simplified-put}).
\begin{lstlisting}[caption={put},label={lst:simplified-put}]
// mixing the client API and system API
put(repl,k,v,ctx) {

    // Dependency for previous reads and writes
    deps = ctx_to_dep(ctx);(*\label{line:put-ctx-to-deps}*)

    // increase local time and appending local kv with a new version.
    inc(repl.local_time);(*\label{line:put-inc-local}*) 
    list_isnert(repl.kv[k],(v, (local_time + id), deps));(*\label{line:put-update-kv}*)

    // update dependency for writes
    ctx.writers += (k,(local_time + id),deps);(*\label{line:put-update-ctx}*)

    // broad case
    asyn_brordcase(k, v, (time ++ id), deps);
    return (local_time + id);
}
\end{lstlisting}

The following is the rule corresponds to \cref{lst:simplified-put}:

\begin{mathpar}
    \inferrule[Put]{%
        ( \stk, \ss, \emptyset ), \pmutate{\key}{\vx} \toTRANS
        ( \stk', \stub, \Set{(\otW, \key, \val )} ), \pskip
        \\\\
        \dep = \Set{(\key', \txid)}[\exsts{i} i \in \viCL(\key') \land \txid = \wtOf(\mkvsCOPS(\key', i)) \lor (\key', \txid) \in \depOf{\mkvsCOPS(\key', i)} ] \texttt{ ---> \cref{lst:simplified-put}, \cref{line:put-ctx-to-deps}} 
        \\\\
        \txid = \min\Set{%
        \txidCOPS{\repl}{\cl}{n'}{0}
        }[%
            \fora{\key', i \in \viREPL(\key'), n} 
            \txidCOPS{\stub}{\stub}{n}{\stub} = \wtOf(\mkvsCOPS(\key',i))
            \implies n' > n 
        ] \texttt{ ---> \cref{lst:simplified-put}, \cref{line:put-inc-local}}
        \\\\
        \mkvsCOPS' = \mkvsCOPS\rmto{\key}{\mkvsCOPS(\key) \lcat (\key, \txid, \emptyset, \dep)} \texttt{ ---> \cref{lst:simplified-put}, \cref{line:put-update-kv}}
        \\\\
        \viREPL' = \viREPL\rmto{\key}{\viREPL(\key) \uplus \Set{\abs{\mkvsCOPS'(\key)} - 1}} \texttt{ ---> \cref{lst:simplified-put}, \cref{line:put-update-kv}}
        \\\\
        \viCL' = \viCL\rmto{\key}{\viREPL(\key) \uplus \Set{\abs{\mkvsCOPS'(\key)} - 1}} \texttt{ ---> \cref{lst:simplified-put}, \cref{line:put-update-ctx}}
    }{%
    \repl, \cl \vdash 
    \mkvsCOPS, \viREPL, \viCL, \stk, \ptrans{\pmutate{\key}{\vx};} \toCMD{\viCL, \Set{(\otW, \key, \val )} }
    \mkvsCOPS', \viREPL', \viCL', \stk', \pskip
    }
\end{mathpar}

The first premiss is to execute the transaction locally (\cref{fig:semantics-trans}).
Since there is only a write, the snapshot \( \ss \) can be any snapshot.
The second line computes the dependency set for the new write operation,
by collecting all the writers of versions included in the view \( \viCL \).
The third line simulates the increment of local time.
Even thought we do not directly track the local time of a replica, 
yet the local time can compute as the maximum time contained in the replica's view \( \viREPL \).
The forth and fifth simulates the updates of the replica's key-value store,
and the last premiss simulates the update of client context.

\paragraph{\bf Read}
The following is a simplified algorithm by directly taking a list of versions 
\verb|ccv| satisfies causal consistency constraint,
\ie the second phase of \cref{lst:cops-client-read},
and then read the versions indicated by \verb|ccv|.
%The simplified algorithm is easier to understand,
%yet it is only for verification purpose because the extremely bad performance.

\begin{lstlisting}[caption={get\_trans},label={lst:get-trans}]
List(Val) get_trans(ks,ctx) {
    // only guarantee to read up-to-date value 
    // the moment reading the individual key
    for k in ks { rst[k] = get_by_version(k,LATEST); } (*\label{line:get-trans-pick-ccv-1}*)

    for k in ks { 
        ccv[k] = max (ccv[k],rst[k].ver); (*\label{line:get-trans-ccv-1}*)
        for dep in rst[k].deps (*\label{line:get-trans-ccv-2}*)
            if ( dep.key (*$\in$*) ks ) ccv[k] = max (ccv[dep.key],dep.vers); (*\label{line:get-trans-ccv-3}*) 
    } (*\label{line:get-trans-ccv-4}*)
    for k in ks   (*\label{line:get-trans-ccv-5}*)
        if ( ccv[k] > rst[k].vers ) rst[k] = get_by_version(k,ccv[k]); (*\label{line:get-trans-ccv-6}*)

    // update the ctx
    for (k,ver,deps) in rst { ctx.readers += (k,ver,deps); } (*\label{line:get-trans-update-ctx-1}*)

    return to_vals(rst); (*\label{line:get-trans-read-1}*)
}                                   
\end{lstlisting}

The following is the rule for read-only transaction:

\begin{mathpar}
    \inferrule[GetTrans]{%
        %\viCL \viewleq \viCL' \viewleq \viREPL  \texttt{ ---> \cref{lst:get-trans}, \cref{line:get-trans-update-ctx-1,line:get-trans-update-ctx-2}}
        \trans =  \plookup{\vx_1}{\key_1}; \dots; \plookup{\vx_j}{\key_j};
        \\
        \vi_0 = \viCL
        \\\\
        {\begin{array}{@{}l@{}}
            \mathsf{for} \ i \ \mathsf{in} \Set{1, \dots, j} \\
            \quad m_i \in \viREPL(\key_i) \\
            \quad \vi'_i = \vi_{i-1}\rmto{\key_i}{\vi_{i-1}(\key_i) \cup \Set{m_i}} 
            \texttt{ ---> \cref{lst:get-trans}, \cref{line:get-trans-pick-ccv-1}} \\
            \quad \vi_i = \lambda \key \ldotp \vi'_i(\key) \cup \Set{x}[(\key,\wtOf[\mkvsCOPS(\key,x)]) \in \depOf[\mkvsCOPS(\key_i,m_i)]]  \\
            \qquad \texttt{ ---> \cref{lst:get-trans}, \cref{line:get-trans-ccv-1,line:get-trans-ccv-2,line:get-trans-ccv-3,line:get-trans-ccv-4,line:get-trans-ccv-5,line:get-trans-ccv-6,line:get-trans-update-ctx-1}} \\
        \end{array}}
        \\\\
        \\\\
        \viCL' = \vi_j
        \\
        ( \stk, \func{getMax}[\mkvsCOPS, \viCL'], \emptyset ), \trans \toTRANS
        ( \stk', \stub, \fp ), \pskip \texttt{ ---> \cref{lst:get-trans}, \cref{line:get-trans-read-1}}
        \\\\
        \txidCOPS{\repl}{\cl}{n'}{n} = \max\Set{\txidCOPS{\repl}{\cl}{z'}{z}}[\txidCOPS{\repl}{\cl}{z'}{z} \in \mkvsCOPS ]
        \\
        \mkvsCOPS' = \updateKV[\mkvsCOPS, \viCL', \fp, \txidCOPS{\repl}{\cl}{n'}{n+1}] 
    }{%
        \repl, \cl \vdash 
        \mkvsCOPS, \viREPL, \viCL, \stk, \ptrans{\plookup{\vx_1}{\key_1}; \dots; \plookup{\vx_j}{\key_j}; } \toCMD{\viCL', \fp}
        \mkvsCOPS', \viREPL, \viCL', \stk', \pskip
    }
    \and
    \inferrule[ClientCommit]{%
        \repl, \cl \vdash 
        \mkvsCOPS, \vienvCOPS(\repl), \vienvCOPS(\cl), \stk, \prog(\cl) \toCMD{\viCL'', \fp}
        \mkvsCOPS', \viREPL', \viCL', \stk', \cmd'
    }{%
        \mkvsCOPS, \vienvCOPS, \thdenv, \prog \toPROG{\viCL'', \fp}
        \mkvsCOPS', \vienvCOPS\rmto{\repl}{\viREPL'}\rmto{\cl}{\viCL'}, \thdenv\rmto{\cl}{\stk'}, \prog\rmto{\cl}{\cmd'}
    }
\end{mathpar}

The for-loop in the premiss picks a version for key \( \key_i \) from \( \viREPL \) and adds to client view, which corresponds to the initial \verb|ccv|\((\key_i)\).
Then it add all the new dependencies \( \depOf[\mkvsCOPS(\key_i,m_i)] \) to the view that yields \( \vi_i \).
The view \( \vi_i \) corresponds to \verb|ccv|\((\key_i)\) after the update against the dependencies.
Given the view \( \viCL \), the client fetches the version with the maximum writer it can observed for each key, which is computed by \( \func{getMax} \) function. 
It is different from \( \snapshot \) because \( \snapshot \) fetches the latest version with respect to the position in the list, but later one we will prove \( \func{getMax} \) and \( \snapshot \) are equivalent.
\begin{align*}
    \func{getMax}[\mkvsCOPS, \viCL] & \defeq 
    \lambda \key \ldotp \left( \max\Set{(\val, \txid, \txidset, \dep)}[\exsts{i} (\val, \txid, \txidset, \dep) = \mkvsCOPS(\key, i) \land i \in \viCL(\key)] \right)\projection{1} \\
    (\val, \txidCOPS{\repl}{\cl}{n}{n'}, \txidset, \dep) & > (\val', \txidCOPS{\repl'}{\cl'}{n''}{n'''}, \txidset', \dep') \defiff (n, \repl, n') > (n'', \repl, n''') 
\end{align*}
The rest part are trivial be picking a new transaction identifier with larger read counter and committing to key-value store.

Last the rule for receiving a update.
A replica updates its local state only if all the dependencies has been receive as shown in \cref{lst:cops-replica-receive-msg}.
\begin{mathpar}
    \inferrule[sync]{%
        \viREPL = \vienvCOPS(\repl)\rmto{\key}{\vienvCOPS(\repl)(\key) \uplus i} 
        \\\\
        {\left(\begin{array}{@{}l@{}}
        \fora{\key', m, \ver} 
        \ver = \mkvsCOPS(\key, i) \land {}
        (\key', \wtOf(\mkvsCOPS(\key', m))) \in \ver\projection{4}
        \implies m \in \viREPL'(\key') 
        \end{array}\right)} 
    }{%
        \mkvsCOPS, \vienvCOPS, \thdenv, \prog \toPROG{}
        \mkvsCOPS, \vienvCOPS\rmto{\repl}{\viREPL}, \thdenv, \prog
    }
\end{mathpar}
The premiss says, the first line, the replica \( \repl \) receive a new version \( i \) for key \( \key \),
and, the second line, only if all the dependencies of the new update already in the replica's view.


\paragraph{\bf COPS key-value store}
To verify the algorithm, we prove that for any COPS trace produced by the algorithm,
there exists a corresponding causal consistency trace.
First we prove that the key-value stores from COPS trace satisfy \cref{def:mkvs}.

\begin{theorem}[Well-formed COPS key-value]
    \label{thm:cops-key-value-well-form}
    Let ignore the dependencies of versions from \( \mkvsCOPS \).
    Given the initial key-value store \( \mkvsCOPS_0 \), initial views \( \vienvCOPS_0 \) and some programs \( \prog_0 \), for any \( \mkvsCOPS_i \) and \( \vienvCOPS_i \)  such that: 
    \[
        \mkvsCOPS_0, \vienvCOPS_0, \thdenv_0, \prog_0 {\toPROG{}}^* \mkvsCOPS_i, \vienvCOPS_i, \thdenv_i, \prog_i
    \]
    The key-value store \( \mkvsCOPS_i \) satisfies \cref{def:mkvs} and any replica or client view \( \vi \) from \( \vienvCOPS_i \) is a valid view of the key-value store, \ie \( \vi \in \Views(\mkvsCOPS_i) \).
\end{theorem}
\begin{proof}
    We need to prove the  \( \mkvsCOPS_i \) satisfies the well-formed conditions,
    and any view \( \vi_i \Views(\mkvsCOPS_i) \).
    We prove it by introduction on the length \( i \).
    \begin{itemize}
    \item \caseB{\(i = 0\)}
        It holds trivially since each key only has the initial version \( (\val_0,\txid_0,\emptyset, \emptyset) \).
        Since there is only the initial version for each key, it is easy to see that any view \( \vi_0 \) satisfying the well-formed conditions in \cref{def:views}.
    \item \caseI{\(i+1\)}
        We perform case analysis on the possible \((i+1)\)-\emph{th} step:
        \begin{itemize}
            \item \rl{Put}
                Assume the client \( \cl \) of a replica \( \repl \) commits a single-write transaction \( \txid \) that installs a new version for key \( \key \).
                By the premiss of \rl{Put}, the new transaction identifier \( \txid = \txidCOPS{\repl}{\cl}{n'}{0} \) where for some \( n' \) that is greater than any \( n \) from any writers \( \txidCOPS{\stub}{\stub}{n}{\stub} \) that are observable by the replica \( \repl \).
                Since the new transaction \( \txid = \txidCOPS{\repl}{\cl}{n'}{0} \) is a single-write transaction which is always installed at the end of the list associated to \( \key \), it is sufficient to prove the following:
                \begin{gather}
                    \fora{j} 0 \leq j < \abs{ \mkvsCOPS_i(\key) } \implies \wtOf(\mkvsCOPS_{i}(\key, j)) \neq \txid \label{equ:write-trans-unique} \\
                    \fora{j, n} \txidCOPS{\repl}{\cl}{n}{\stub} \in \Set{\wtOf(\mkvsCOPS_{i}(\key,j))} \cup \rsOf(\mkvsCOPS_{i}(\key, j)) \implies n < n' \label{equ:replica-time-monotonic-inc}
                \end{gather}
                \cref{lem:mono-local-time} implies \cref{equ:write-trans-unique} and \cref{equ:replica-time-monotonic-inc}.
                Thus the new key-value store \( \mkvsCOPS_{i+1} \) satisfies the well-formed conditions.
                Now let consider the views, especially the views of the replica \( \viREPL' \) and the client \( \viCL' \).
                Because views different replicas or clients remain unchanged, by \ih they satisfy \( \vi' \in \Views(\mkvsCOPS_{i+1}) \).
                The new view for replica \( \viREPL' = \viREPL\rmto{\key}{\abs{\mkvsCOPS_{i+1}(\key)} - 1} \)
                where \( \viREPL \) is the replica's view before updating and the writer of the last version of \( \key \) is \( \txid \).
                Because \( \txid \) is a single-write transaction, so the new view \( \viREPL' \) still satisfies the atomic read.
                For similar reason, the new view for client \( \viCL' \) till satisfies atomic read.
                Therefore we have \( \viREPL', \viCL' \in \Views(\mkvsCOPS_{i+1}) \).

            \item \rl{GetTrans}
                Assume the client \( \cl \) of a replica \( \repl \) commits a read-only transaction \( \txid \).
                Since it is a read-only transaction, it suffice to prove the following:
                \begin{gather}
                    \fora{\key,j} 0 \leq j < \abs{ \mkvsCOPS_i(\key) } \implies \txid \notin \rsOf(\mkvsCOPS_{i}(\key, j)) \label{equ:read-trans-unique} 
                \end{gather}
                \cref{lem:mono-local-time} implies \cref{equ:read-trans-unique}.
                Thus the new key-value store \( \mkvsCOPS_{i+1} \) satisfies the well-formed conditions.
                Now let consider the views.
                Since only the client view has changed, it is sufficient to  prove that \( \viCL' \in \Views(\mkvsCOPS_{i+1}) \), 
                where \( \viCL' \) is the new client view.
                By \cref{lem:write-unique}, the new view \( \viCL' \) satisfies the atomic read constraint.
                Therefore \( \viCL' \in \Views(\mkvsCOPS_{i+1})  \).
        \end{itemize}
    \end{itemize}
\end{proof}



\begin{lemma}[Monotonic local time]
    \label{lem:mono-local-time}
    Given a reduction step such that: 
    \[
        \mkvsCOPS, \viREPL, \viCL, \stk, \cmd {\toCMD{}}^* \mkvsCOPS', \viREPL', \viCL', \stk', \cmd'
    \]
    let \( \txidCOPS{\repl}{\cl}{n}{n'} \) be the new transaction, \ie \( \txidCOPS{\repl}{\cl}{n}{n'} \in \mkvsCOPS' \land  \txidCOPS{\repl}{\cl}{n}{n'} \notin \mkvsCOPS \).
    It implies the new transaction is greater than any transaction committed by the same client view \( \viCL \), \ie
    \[ 
        \txidCOPS{\repl}{\cl}{n''}{n'''} \in \mkvsCOPS \implies (n,\repl,n') > (n'', \repl, n''')
    \]
\end{lemma}
\begin{proof}
    We perform case analysis.
    \begin{itemize}
        \item \rl{Put}
            By the premiss of the rule, 
            the new transaction is in the form \( \txidCOPS{\repl}{\cl}{n}{0} \).
            By the \cref{lem:repl-observe-own} we have for any existing transaction \( \txidCOPS{\repl}{\cl}{n'}{\stub} \),
            \[
                \txidCOPS{\repl}{\cl}{n'}{\stub} \in \mkvsCOPS \implies n > n'
            \]
        \item \rl{GetTrans}
            Let \( \txidCOPS{\repl}{\cl}{n}{n'} \) be the new transaction.
            By the premiss we have that for any existing transaction \( \txidCOPS{\repl}{\cl}{n''}{n'''} \), 
            \[
                \txidCOPS{\repl}{\cl}{n''}{n'''} \in \mkvsCOPS \implies n = n'' \land n' > n'''
            \]
    \end{itemize}
\end{proof}

\begin{lemma}[Unique writer]
    \label{lem:write-unique}
    Each version has a unique writer.
\end{lemma}
\begin{proof}
    Because a transaction can write to at most one key, and 
    a client at least observes its own writes (the premiss of \rl{Put})
    we have the proof.
\end{proof}

\paragraph{\bf COPS normal trace}
We define COPS semi-normal trace then normal trace.
Later, we prove for any COPS trace, there exists an equivalent normal trace.

The dependency of a transaction \( \depOf[\mkvsCOPS, \txid] \) is defined as:
\begin{itemize}
    \item if \( \txid \) is a single-write transaction:
    \begin{align*}
        \depOf[\mkvsCOPS, \txid] & \defeq 
        \begin{multlined}
            \dep  \ \text{where} \ 
            \exsts{\key,j} \lastConf(\tr)\projection{1}(\key,j) = (\stub, \txid, \stub, \dep)
        \end{multlined}
    \end{align*}
    \item if \( \txid \) is a read-only transaction:
    \begin{align*}
        \depOf[\mkvsCOPS, \txid] & \defeq 
        \begin{multlined}
            \bigcup_{\Set{\ver}[\exsts{\key,j} \ver = \mkvsCOPS(\key,j) \land \txid \in \rsOf(\ver)]} \depOf[\ver] \\
        \end{multlined}
    \end{align*}
\end{itemize}
Given a trace \( \tr \), 
the function \( \func{maxVersion}[\tr, \txid] \) returns the version that the transaction \( \txid \) depends on
and is the last one that appears in the trace:
\begin{align*}
    \func{maxVersion}[\tr, \txid] & \defeq \txid_i
    \quad \begin{multlined}[t]
        \text{the \( \txid_i \) is the last one appears in \( \tr \) such that} \\
        (\key',\txid_i) \in \depOf[\lastConf(\tr)\projection{1}, \txid] \land \txid_i \notin \mkvsCOPS_i \land  \txid_i \in \mkvsCOPS_{i+1}
    \end{multlined}
\end{align*}
Give two COPS's traces \( \tr \) and \( \tr' \), \( \mkvsCOPS \) being the final state of \( \tr \) and \( \mkvsCOPS' \) for \(\tr'\),
if the two traces are equivalent, if and only if,
\[ 
    \begin{array}{@{}l@{}}
    \fora{\txid,\fp,\dep} \txid \in \mkvsCOPS
    \land \fp = \func{RWset}[\mkvsCOPS,\txid] \land \dep =  \depOf[\mkvsCOPS,\txid] \\
    \quad {} \iff
    \txid \in \mkvsCOPS'
    \land \fp = \func{RWset}[\mkvsCOPS',\txid] \land \dep =  \depOf[\mkvsCOPS',\txid] \\
    \end{array}
\]
where \( \func{RWset}[\mkvsCOPS,\txid] \) is the read-write set of \( \txid \).
Note that \( \func{RWset} \) is well-defined by \cref{thm:cops-key-value-well-form}.

A COPS's semi-normal trace is a trace \( \tr \) if it is in the form that
read-only transactions \( \txid_\rd \) directly follows its \( \func{maxVersion}[\tr, \txid_\rd] \) or another read-only transaction \( \txid_{\rd}' \) such that \( \func{maxVersion}[\tr, \txid_\rd]  =  \func{maxVersion}[\tr, \txid_{\rd}']\).


\begin{corollary}
    \label{lem:cops-semi-normal-trace}
    \label{cor:cops-semi-normal-trace}
    For any COPS's trace \( \tr \), there exists a equivalent semi-normal trace \( \tr' \) 
    such that \( \lastConf(\tr) = \lastConf(\tr') \).
\end{corollary}
\begin{proof}
    It is easy to prove by induction on the numbers of read-only transactions that are not in the wanted position.
    Let take the first one for those read-only transactions \( \txid_\rd \) who does not follows its \( \func{maxVersion}[\tr, \txid_i] \).
    It is safe to move the reduction step to the right in the trace,
    until it directly follows its \( \func{maxVersion}[\tr, \txid_\rd] \) or another read-only transaction \( \txid_{\rd}' \) such that \( \func{maxVersion}[\tr, \txid_\rd] = \func{maxVersion}[\tr, \txid_{\rd}']\). 
    Assume it is \(i\)-\emph{th} step, and assume  \( (i-1)\)-\emph{th} is not  \( \func{maxVersion}[\tr, \txid_\rd] \) nor another read-only transaction \( \txid_{\rd}' \) such that \( \func{maxVersion}[\tr, \txid_\rd] = \func{maxVersion}[\tr, \txid_{\rd}']\).
    It means the \( (i-1)\)-\emph{th} step must be a write that \( \txid_\rd \) does not depend on.
    Because if \( (i-1)\)-\emph{th} step is a read only transaction \( \txid''_\rd \) 
    it must be the case 
    \( \func{maxVersion}[\tr, \txid_\rd] = \func{maxVersion}[\tr, \txid_{\rd}'']\), 
    otherwise \( \txid_\rd \) is not the first read-only transaction that is not in the wanted place.
\end{proof}

A COPS's normal trace is a trace \( \tr \) if it is semi-normal trace and the single-write transactions step appears in the trace in the order of the writers of those transactions.

\begin{theorem}[normal trace]
    \label{thm:cops-normal-trace}
    \label{lem:cops-normal-trace}
    \label{cor:cops-normal-trace}
    For any COPS's trace \( \tr \), there exists an equivalent normal trace \( \tr' \).
\end{theorem}
\begin{proof}
    By \cref{cor:cops-semi-normal-trace}, let consider a semi-normal trace \( \tr \).
    It is easy to prove by induction on the single-write transactions that are out of order.
    Let take the first single-write transaction \( \txid = \txidCOPS{\repl}{\cl}{n}{n'} \) that is out of order.
    Suppose it is the i-\emph{step}.
    Assume \( \txid \) write to key \( \key \) with value \( \val \).
    Assume the preview write-only transaction \( \txid'  = \txidCOPS{\repl'}{\cl'}{n''}{n'''} \) and it follows \( (n,\repl,n') < (n'', \repl', n''') \).
    The intuition here is to swap the reduction steps of these two transactions and those read-only transactions following them.
    We assume \( \txid \)  and \( \txid' \) write to the same key, otherwise it is safe to swap.
    For any read-only transactions \( \txid_{\rd} \in \txidset_{\rd} \) following \( \txid \), we knows \( \txid' \notin \depOf[\mkvsCOPS_i, \txid_{\rd}] \), otherwise it violate \( \func{maxVersion}[\tr, \txid_\rd] = \txid \).
    Thus it is safe to swap the reduction steps and alter views that includes either or both versions.
    Note that if a view \( \vi \) includes both \( \txid \)  and \( \txid' \), and if \( \txid' \) also write to the same key \( \key \) with value \( \val' \), the \( \vi \) will fetch the value from \( \txid' \):
    \begin{equation}
        \label{equ:out-order-wrte-older-value}
        \func{getMax}[\mkvsCOPS_i,\vi](\key) = \val'
    \end{equation}
\end{proof}

\begin{corollary}
    \label{cor:get-max-to-get-snap}
    For the \rl{Put} and \rl{GetTrans}, it is safe to replace the function \( \func{getMax} \) with the function \( \snapshot \).
\end{corollary}
\begin{proof}
    By \cref{thm:cops-normal-trace}, each trace \( \tr \) has an equivalent normal trace \( \tr' \).
    Assume i-\emph{th} step in the trace \( \tr' \).
    Assume the key-value store before is \( \mkvsCOPS_i \).
    For any versions \( m \) and \( j \) from the same key in the \( \tr' \):
    \[
        0 < m < j < \abs{\mkvsCOPS_i(\key)} \implies \wtOf(\mkvsCOPS_i(\key,m)) <  \wtOf(\mkvsCOPS_i(\key,j))
    \]
    thus it is safe to use \( \snapshot \).
\end{proof}

\paragraph{\bf COPS normal trace to \( \ET_\CC \) trace}
Given \cref{thm:cops-key-value-well-form,thm:cops-normal-trace,cor:get-max-to-get-snap}, for any COPS's trace \( \tr \), 
there exists a trace \( \tr' \) that is a normal trace and satisfies \( \ET_\top \).
Then by the following theorem (\cref{thm:cops-cc}), we proof the COPS trace satisfies \( \ET_\CC \).

\begin{theorem}[COPS satisfying causal consistency]
    \label{thm:cops-cc}
    Given a trace starting from the initial key-value store \( \mkvsCOPS_0 \), initial views \( \vienvCOPS_0 \) and some programs \( \prog_0 \), for any \( \mkvsCOPS_i \) and \( \vienvCOPS_i \)  such that: 
    \[
        \mkvsCOPS_i, \vienvCOPS_i, \thdenv_i, \prog_i \toPROG{\viCL,\fp} \mkvsCOPS_{i+1}, \vienvCOPS_{i+1}, \thdenv_{i+1}, \prog_{i+1} 
    \]
    then the i-\emph{th} step satisfies \( \ET_\CC \), \ie
    \[
        \ET_\CC \vdash ( \mkvsCOPS_i, \viCL ) \csat \fp : (\mkvsCOPS_{i+1},\vienvCOPS_{i+1}(\cl) )
    \]
\end{theorem}
\begin{proof}
    We prove this by induction on the length of trace \( i \).
    We introduce two invariants \( I_1, I_2 \).
    The \( I_1 \) states that for any view \( \vi \), if the view \( \vi \) includes a version, 
    it also includes all the version it depends on, that is,
    \[
        \begin{array}{@{}l@{}}
        \fora{\cl, \vi,\key,\key',j,m,\dep} \\
        \quad \vi = \func{co-dom}[\vienvCOPS_i] \land j \in \viCL(\key') 
        \land (\stub,\stub,\stub,\dep) = \mkvsCOPS_i(\key,j)
        \land (\key',\wtOf(\mkvsCOPS_i(\key',m))) \in \dep \\
        \qquad \implies m \in \vi(\key')
        \end{array}
    \]
    The \( I_2 \) states that for any transaction \( \txid \), 
    \( \depOf[\mkvsCOPS_i,\txid] \) includes transactions that \( \txid \) depends on
    with respect to \( (\SO \cup \WR) \).
    \[
        \begin{array}{@{}l@{}}
        \fora{\txid} 
        \txid \in \mkvsCOPS 
        \implies 
        ((\SO \cup \WR)^{-1})^*(\txid) \subseteq \depOf[\mkvsCOPS_i,\txid]
        \end{array}
    \]
    if the view \( \vi \) includes a version, 
    Since \( \ET_\CC = \ET_\MR \cap \ET_\RYW \cap \allowed[\SO \cup \WR] \), we prove the three constraints separately.
    In each case we need to consider \rl{Put}, \rl{GetTrans} and \rl{Sync}
    Let \( \txid \) be the new transaction committed to replica \( \repl \) by the client \( \cl \).
    Let \( \viCL \) and \( \viCL' \) be the client views before and after respectively,
    and \( \viREPL \) and \( \viREPL' \) be the replica views before and after respectively.
    \begin{itemize}
        \item \( \ET_\MR \). 
            For \( \rl{Put} \), it is easy to see \( \viCL \viewleq \viCL' \) and \( \viREPL \viewleq \viREPL' \).
            For \( \rl{GetTrans} \), it is known that \( \viREPL = \viREPL' \). 
            In the premiss of \( \rl{GetTrans} \),
            it fetches a version, \( m_i \)-\emph{th} version, from the replica for each key \( \key_i \) and adds all dependencies, which are associated with version, in the client view.
            This means \( \viCL \viewleq \viCL' \).

        \item \( \ET_\RYW \). 
            By \( \ET_\MR \), it suffices to only consider 
            if the newly committed transaction is included in the view.
            We only need to consider \rl{Put}.
            The new views \( \viCL' = \viCL\rmto{\key}{j} \) and \( \viREPL' = \viREPL\rmto{\key}{j} \)
            where the new version \( \mkvsCOPS_{i+1}(\key,j)\) is written by the client \( \cl \).
            
        \item \( \allowed[\SO \cup \WR] \).
            Given the invariants \( I_1 \) and \( I_2 \),
            let consider \rl{Put}, \rl{GetTrans} and \rl{Sync}.
            \begin{itemize}
                \item \rl{Put}.
                By invariants \( I_1 \) and \( I_2 \),
                we know \( \viCL = \getView[\mkvsCOPS_i,((\SO \cup \WR)^{-1})^{*}(\Tx[\mkvsCOPS_i,\viCL])]\).
                Now we need to re-establish the \( I_1 \) for the new views \( \viCL' \) and \( \vienvCOPS' \) and \( I_2 \) for the new transaction \( \txid \).
                Since it is a single-write transaction,
                assume it appends \( j\)-\emph{th} version to key \( \key \).
                This means the new client views \( \viCL' = \viCL\rmto{\key}{ \viCL(\key) \cup \Set{j} } \),
                Let \( \dep \) be the dependency for the new version, \ie \( \dep = \depOf(\mkvsCOPS_{i+1}(\key,j)) \).
                By the premiss, the \( \dep \) includes all versions included in \( \viCL \).
                This means for any key \( \key' \) and index \( m \):
                \[
                    (\key',\wtOf[\mkvsCOPS_i(\key',m)]) \in \dep \implies m \in \viCL'(\key')
                \]
                Since \( \viCL \viewleq \vienvCOPS \),
                therefore the invariant \( I_1 \) has been preserved.
                \item 
                For \rl{GetTrans}, the client \( \cl \) commits a read-only transaction.
                By the premiss of the rule, the new view \( \vienvCOPS_{i+1}(\cl) \) satisfies the invariant.
            \end{itemize}

        \item \( \ET_\WFR \).
            By \cref{lem:cops-dep} and the invariant, we have to proof.
            Now we need to re-establish the invariant \( I_{i+1} \).
            \begin{itemize}
                \item 
                For \rl{GetTrans}, the client \( \cl \) commits a read-only transaction.
                By the premiss of the rule, the new view \( \vienvCOPS_{i+1}(\cl) \) satisfies the invariant.
                \item For \rl{Put}, the new view \( \vienvCOPS_{i+1}(\cl) = \viCL\rmto{\key}{j} \),
                where the new version \( \mkvsCOPS_{i+1}(\key,j)\) is written by the client \( \cl \).
                Let \( \dep \) be the dependency for the new view, \ie \( \dep = \depOf(\mkvsCOPS_{i+1}(\key,j)) \).
                By the premiss, the \( \dep \) includes all version included in \( \viCL \), 
                therefore the invariant has been preserved.
            \end{itemize}
    \end{itemize}
\end{proof}

\begin{lemma}
    \label{lem:cops-dep}
    Given the initial key-value store \( \mkvsCOPS_0 \), initial views \( \vienvCOPS_0 \) and some programs \( \prog_0 \), for any \( \mkvsCOPS_i \) and \( \vienvCOPS_i \)  such that: 
    \[
        \mkvsCOPS_0, \vienvCOPS_0, \thdenv_0, \prog_0 {\toPROG{}}^* \mkvsCOPS_i, \vienvCOPS_i, \thdenv_i, \prog_i 
    \]
    for any versions \( ( \val,\txid,\txidset,\dep ) = \mkvsCOPS_i(\key,j) \) and \( ( \val',\txid',\txidset',\dep' ) = \mkvsCOPS_i(\key',m) \), if \( \txid \) depends on \( \txid' \), \ie
    \[
        \begin{array}{@{}l@{}}
            \txid' \toEDGE{\SO\rflx} \txid
            \lor \exsts{\txid''}
            \txid'' \in \txidset' \land
            \txid'' \toEDGE{\SO\rflx} \txid
        \end{array}
    \]
    then \( \txid' \) and \( \dep' \) is in the dependency \( \dep \), \ie
    \[
        (\stub, \txid') \in \dep \land \dep' \subseteq \dep
    \]
\end{lemma}
\begin{proof}
    We prove by induction on the length of trace \( i \).
    \begin{itemize}
        \item 
        \caseB{\(i = 0\)}
        It trivially holds.
        \item  
        \caseI{\( i > 0 \)}
        Suppose \( \mkvsCOPS_i \) satisfies the property, let consider the next key-value store \( \mkvsCOPS_{i+1}\).
        We perform case analysis.
        \begin{itemize}
            \item 
            \rl{Put}.
            Assume a client \( \cl \) commit a new transaction \( \txid \) that installs a new version to key \( \key \).
            Assume the initial view for the client is \( \viCL \).
            In the new key-value store \( \mkvsCOPS_{i+1} \), for versions other than \( \mkvsCOPS_{i+1}(\key, \abs{\mkvsCOPS_{i+1}} - 1 ) \), they still satisfy the property by \ih
            By \cref{lem:cl-observe-own}, it follows that for any preview transaction from the same client \( \txid' \):
            \[
                \begin{array}{@{}l@{}}
                    \txid' \toEDGE{\SO\rflx} \txid
                    \lor \exsts{\txid''}
                    \txid'' \in \txidset' \land
                    \txid'' \toEDGE{\SO\rflx} \txid
                \end{array}
            \]
            the version written by the transaction \( \txid' \) is included in the view \( \viCL \).
            By the premiss of \rl{Put} that how the new \( \dep \) is constructed, we have proof.

            \item 
            \rl{GetTrans}.
            Since it only commits a read-only transaction, 
            the \( \mkvsCOPS_{i+1} \)  satisfies the property by \ih
        \end{itemize}
    \end{itemize}
\end{proof}

%\begin{theorem}[COPS satisfy parallel snapshot isolation]
    %For any COPS's trace \( \tr \) starting from the initial key-value store \( \mkvsCOPS_0 \), initial views \( \vienvCOPS_0 \) and some programs \( \prog_0 \), 
    %there exists an equivalent trace \( \tr' \) where \( \mkvsCOPS'_i \) and \( \vienvCOPS'_i \)  such that: 
    %\[
        %\mkvsCOPS'_i, \vienvCOPS'_i, \thdenv'_i, \prog'_i \toPROG{\viCL',\fp'} \mkvsCOPS'_{i+1}, \vienvCOPS'_{i+1}, \thdenv'_{i+1}, \prog'_{i+1} 
    %\]
    %then the i-\emph{th} step satisfies \( \ET_\PSI \), \ie
    %\[
%\ET_\PSI \vdash ( \mkvsCOPS'_i, \viCL' ) \csat \fp' : ( \mkvsCOPS'_{i+1}, \vienvCOPS'_{i+1}(\cl) )
    %\]
%\end{theorem}
%\begin{proof}
    %By \cref{thm:cops-cc}, we already the trace \( \tr \) satisfying \( \ET_\CC \).
    %It is sufficient to construct the trace \( \tr' \) so that 
%\( \ET_\PSI \vdash ( \mkvsCOPS'_i, \viCL' ) \csat \fp' : ( \mkvsCOPS'_{i+1}, \vienvCOPS'_{i+1}(\cl) ) \) since \( \ET_\PSI = \ET_\CC \cap \ET_\UA\).
    %We inductively alter all single-write transactions \( \txid \) that 
    %does not start with a view containing all the versions for the key the transactions \( \txid \) want to write.
    %Initially let \( \tr' = \tr \).
    %Let pick the first single-write transaction \( \txid \), assuming it is i-\emph{step}, 
    %who installs a version \( \mkvsCOPS_{i+1}(\key,j) = (\val,\txid,\txidset,\dep )\) with a initial view \( \viCL \) does not contain all versions \( \mkvsCOPS_i(\key,m) \)
    %for some key \( \key \), client \( \cl \) and index \( m \) and \( j \).
    %Since \( \txid \) is the first transaction does not contain all versions \( \mkvsCOPS_i(\key,m) \),
    %it means the view \( \viCL \) does not include only the last version \( \mkvsCOPS_i(\key,j-1) \).
    %Suppose the version \( \mkvsCOPS_i(\key,j-1) = (\val',\txid',\txidset',\dep' ) \) for some transaction \( \txid' \).
    %Note that \( \txid' \)  and \( \txid \) comes from different client, otherwise it violate \( \ET_\RYW \).
    %Thus let assume the initial view for \( \txid' \) is \( \viCL[\cl'] \) for client \( \cl' \) such that \( \cl' \neq \cl \).
    %We performance case analysis as the following.
    %\begin{itemize}                                                               
        %\item Suppose \( \depOf(\txid') \subseteq \depOf(\txid_\rd) \), for all read-only transactions \( \txid_\rd \in \txidset \).
            %It means it is safe to include \( j-1 \) version into \( \viCL \),
            %\ie the new initial view for \( \txid \) is \( \viCL' = \viCL\rmto{\key}{j-1} \).
            %Given that, all the initial views for the read-only transactions \( \txid_\rd \in \txidset \) also include \( \mkvsCOPS(\key,j) \).
            %Yet because \( \depOf(\txid') \subseteq \depOf(\txid_\rd) \), the version \(\mkvsCOPS(\key,j)  \) is already in the initial view of \( \txid_\rd \).
            %By the new initial view \( \viCL' \), it is easy to see 
            %\( \ET_\UA \vdash (\mkvsCOPS_i, \viCL') \csat \Set{(\otW, \key, \val)}, \viCL'' \).
        %\item there exists a read-only transaction \( \txid_\rd \in \txidset \) such that 
              %\( \depOf(\txid') \not\subseteq \depOf(\txid_\rd) \).
              %\begin{itemize}
                  %\item If it is safe to include \( \depOf(\txid') \) into \( \depOf(\txid_\rd) \), consequently changing the initial for \( \txid_\rd \), for all read-only transactions \( \txid_\rd \in \txidset \), then we are back to the first case.
                  %\item If there exists a read-only transaction \( \txid_\rd \) that is not safe to include \( \depOf(\txid') \).
                      %That is, there exists a single-write transaction  \( \txid'' \) who installs a version \( \mkvsCOPS_i(\key',x)\) for a key \( \key' \) and a index \( x \) and the \( \txid_\rd \) read this version.
                      %Given that it is not safe for  \( \depOf(\txid_\rd) \) to includes \( \depOf(\txid') \), 
                      %there exists anther version \( \mkvsCOPS_i(\key',y) \) for the same key \( \key' \) and a index \( y \) such that \( y > x \).
                      %It follows 
                      %\[ 
                        %\depOf{\mkvsCOPS_i(\key',x)} \subseteq \depOf{\mkvsCOPS_i(\key',y) \subseteq \depOf{\txid'} } 
                      %\]
                      %therefore 
                      %\[
                          %\fora{\txid'_\rd \in \txidset'} \depOf{\mkvsCOPS_i(\key',x)} \subseteq \depOf{ \txid'_\rd}
                      %\]
            %\end{itemize}
    %\end{itemize}
%\end{proof}
